---
title: "EDA-Assignment 1 - Ram Yoogesh"
output: html_notebook
---

## Question 3
## Loading the CSV data 

```{r}
path_concat <- function(path1, path2, sep="/") {
paste(path1, path2, sep = sep)
}
dataDirectory = "/home/yoogesh/Desktop/stat847assignments/data/"
data <- read.csv(file = path_concat(dataDirectory, "class_data.csv"))

```

## Previwing the content

```{r}
#View(data)
data
str(data)
names(data)
```

## a)

$E(D)$ $=$ $(1/10)$ $*$ $(0+1+2+3+4+5+6+7+8+9)$ $=$ $4.5$

## b)

$E(D^2)$ $=$ $\frac{1}{10}$ $*$ $(0^2+1^2+2^2+3^2+4^2+5^2+6^2+7^2+8^2+9^2)$ $=$ $28.5$

$Var(D)$ $=$ $E(D^2)$ $-$ $E(D)^2$ $=$ $28.5$ $-$ $20.25$ $=$ $8.25$

$SD(D)$ $=$ $\sqrt(Var(D))$ $=$ $2.872$
    
## c)
### Median is 4.5

## d)
## (i)
### Binom Dist

## (ii)
$pr(X = x)$ = $\binom{n}{x}$ $(\frac{1}{10})^x$ $*$ $(\frac{9}{10})^{n-x}$ 
        
### (iii)
$E(X)$ = $n$ $*$ $\frac{1}{10}$


## e)
## (i)
```{r}
mean(data$random_digit)
mean(data$student_digit)
```

## (ii)
```{r}
sd(data$random_digit)
sd(data$student_digit)
```

## (iii)
```{r}
median(data$random_digit)
median(data$student_digit)
```

## (iv)
### RD (Random Digit) 
### Mean > D
### S.D = D
### Median > D

### SD (Student Digit)
### Mean < D
### S.D = D
### Median > D

## (v)
### Mean and S.D of RD is close to D 
### Median of SD is close to D


## f)
### when x = 0 
$\binom{42}{0}$ $*$ $(1/10)^0$ $*$ $(9/10)^{42}$ $=$ $0.01197$ 
### When x = 5 
$\binom{42}{5}$ $*$ $(1/10)^5$ $*$ $(9/10)^{37}$ = $0.1724$
### When x = 10
$\binom{42}{10}$ $*$ $(1/10)^{10}$ $*$ $(9/10)^{32}$ = $0.0050$



## 2)

## a)
## (i)
```{r}
stem(data$student_digit)
```

## (ii)
```{r}
stem(data$random_digit)
```

## (iii)
```{r}
rand<-c(0:9)
tot <- sample(rand, length(data$random_digit), replace = TRUE)
stem(tot)
```
### Student_digit looks more like it might have come from Uniform on the digits


## b)

## (i)

```{r}
my_digits<-c(0,1,3,4,7,1,4,9,7,4)
count_digits<-function(d){
  fl<-c()
  n<-length(d)
  for(i in 1:10)
    {
    nv<-0
    for(j in 1:n)
      {
      if(d[j]==i-1)
        {
        nv<-nv+1}  
      }
    fl[i]<-nv
  }
  fl
}
count_digits(my_digits)
```

## (ii)
```{r}
count_digits(data$student_digit)
count_digits(data$random_digit)
```

## (iii)
```{r}
Pearson_chi_sq <- function(observed,expected){
  expected <- sum(observed)/length(observed)
  vect<-c()
  dum<-c()
  n<-length(observed)
  dum<-0
  for(i in 1:n){
    vect[i]<-expected
  }
  for(i in 1:n){
    dum <- dum+((observed[i]-expected)^2)/expected
  }
 return(dum)
}
Pearson_chi_sq(count_digits(data$student_digit))
```

## (iv)
```{r}
## function for student digit
Pearson_chi_sq(count_digits(data$student_digit))
chisq.test(count_digits(data$student_digit))$statistic

## function for random digit
Pearson_chi_sq(count_digits(data$random_digit))
chisq.test(count_digits(data$random_digit))$statistic
```
### Values from Pearson_chi_sq and chisq.test match


## (v)
```{r}
pchisq(Pearson_chi_sq(count_digits(data$student_digit)), 9, lower=FALSE)
chisq.test(count_digits(data$student_digit))$p.value
pchisq(Pearson_chi_sq(count_digits(data$random_digit)), 9, lower=FALSE)
chisq.test(count_digits(data$random_digit))$p.value
```
### The values from pchisq using pearsonchisq matches with chisq.test


## (vi)
```{r}
get_chisqs<-function(n, B=1000){
  rand <- c(0:9)
  vect <- c()
  nv <- c()
  for(i in 1:B){
   tot <- sample(rand ,n, replace=TRUE)
   cs <-count_digits(tot)
   nv[[i]]<-cs
  }
  val <- sapply(nv,Pearson_chi_sq) 
   print(val)
}
n<-nrow(data)
results<-get_chisqs(n=n,B=1000)
```

## (vii)
```{r eval=TRUE}
n <- nrow(data)
B <- 10000
set.seed(314159)

chisq_stats<-get_chisqs(n=n, B=B)
hist(chisq_stats, col="lightgrey", main="Simulated Pearson test null distribution", xlab="test stat")
abline( v=Pearson_chi_sq( count_digits( data$student_digit)), lwd=3, lty=2, col="red")
abline( v=Pearson_chi_sq( count_digits( data$random_digit)), lwd=3, lty=2, col="blue")
```
### Radom digits seem less likely to have been generated as a random sample from a uniform distribution of digits. The main reason behind this is that it has a lower val as we can infer it from the hist. 

## (viii)
```{r}
mean(chisq_stats >= Pearson_chi_sq( count_digits( data$student_digit)))
mean(chisq_stats >= Pearson_chi_sq( count_digits( data$random_digit)))
```

## (ix)
### Since the p value is large for student digit and small for random digit, we can say that the hypothesis is true for student digit anf false for random digit.

