---
title: "Assignment 2 - Ram Yoogesh (20867060)"
output: html_notebook
---

## Question 1 

## (c) (i)
### Writing the r_unifgenFX function to return random derivates 
```{r}
r_unifgenFX <- function(n, qfunction = qnorm)
{
  randomderivates <- runif(n, min =0, max = 1)
  return (randomderivates)
}
```

## (c) (ii)
### Executing the following snippets as instructued 
```{r}
set.seed(1234567)

oldpar <- par(mfrow = c(1, 2))
hist(r_unifgenFX(1000))

hist(r_unifgenFX(1000, qfunction = ruinf))
par(oldpar)
```
## (c)(iii)
### Generating pseudo-random observations from a student t distribution 
```{r}

r_unifgen <- function (n, dof)
{
  robs_student_t <- rt(n, df=dof)
  return (robs_student_t)
}
```

```{r}
sam_obs <- r_unifgen(1000, 3)

# Now passing the vector of random values to the qt fucntion to find the quantile values. 

 x_val <- qt(sam_obs, df= 3, lower.tail = TRUE)

hist(qt(sam_obs, df = 3))
 
``` 

## (d)
```{r}
# Just analyzing the dataset before interpreting the questions ! 

str(mtcars)
# we can infer that it has 11 variables and 32 rows . Names of these are mpg, cyl, disp, hp, drat, wt, qsec, vs, am, gear, card
head(mtcars)

```


## (d)(i)
## In order to answer this question ,lets use the summary function to get some insights. We can infer from summary of qsec ( min value- 14.50, 1st quartile - 16.89 and so on). 
```{r}
summary(mtcars)
```
```{r}
# Now lets use the Quantile function to answer the question 
quantile(mtcars$qsec)
```
## Qsec - The variable denotes the 1/4 mile time took by cars.
## Qunatile function basically prodcues the values at the respective quartiles. Say 0, 25, 50, 75 and 100. What it means for Qsec is that at 25th pecent quartile the value of qsec is 16.89 (It tooks a car 16.89s to reach 1/4 mile). At 50th percent, also called as the median of the dataset, the value at 50th percent quartile is 17.71s, likewise for the remaining quartiles.


## d(ii)
## we can pass mtcars$qsec in the X argument and for the probability we can use runif to generate 1000 observations
```{r}
samobs_fromqsec <- quantile(mtcars$qsec, prob = runif(1000))
samobs_fromqsec
```


## (d)(iii)
## Before answering the question, let us first analyse the cylinders data from mtcars dataset.
```{r}
# Printing all 32 rows of data with respect to number of cylinders
print(mtcars$cyl)

# Running the summary to get more info
summary(mtcars$cyl)

#calculating the mode of cyl
y <- mtcars$cyl
y <- table(y)
names(y)[which.max(y==max(y))]
```
```{r}
## Using the quantile function for cyl
quantile(mtcars$cyl, prob=runif(10))
```

## Yes, it does work for cyl. From the actual data cyl has only 3 varieties (namely 4, 6 and 8 ). So when we compute quantile, we will get these 3 values irrespective of the probabilities we choose. 



## d(iv)
```{r}
dat <- par(mfrow = c(1,2))
hist(mtcars$qsec, main = "Qsec from Mtcars", xlab = "Qsec - (1/4 mile Time)", col = 'red')
hist(samobs_fromqsec, main = "Qsec Generated from Quantile", xlab = "Qsec - (1/4 Time) - Generated", col='skyblue')
par(dat)

```

## From the histograms we can infer that both the graphs follow the same distribution. Although in the left (Original Qsec data from mtcars) we can see that there is no bin at 21-22 seconds - That is, there was no car that took 21-22 seconds to attain 1/4 a mile. On the contrary on the right (Qsec generated using the quantile function) we can see that there is a bin at 21-22 since the data was geneated with prob 0 and 1. On the Y-Axis (Left side Graph) we can see that the frequency goes from 0 to 10 with a step size of 2. Whereas on the Y-axis (Right side graph) we can see that the frequency goes from 0 to 250 with a step size of 50.


## d(v)
```{r}
qqplot(mtcars$qsec, samobs_fromqsec, main = "Q-Q Plot",xlab = "Theoretical Observations", ylab = "Sample Observations")

```

## Generally QQ plots a scatterplot using two quantiles against each other.From the graph we can infer that the points form a line that's roughly straight. This states that both the quantiles come from the same distribution. 

## d(vi)
## Step 1 - Inorder to produce a bootstrap distribution for some estimator theta, Let us use quantile to produce some data. I am using cylinders from mtcars dataset. I used the quantile fucntion with number of bootstrap samples drawn as 100 with a two sided confidence interval(95%) argument. 
```{r}

quantile(mtcars$cyl, probs = runif(50), CI.type = "two.sided", nbboot=100)

```
## I would recommend using quantiles over bootsrtap. The main reason behind is that quantile function is computationally less expensive when compared to bootstrapping. 


