---
title: "A5 (Q3) - Ram Yoogesh Gopu (20867060)"
output:
  pdf_document: default
  html_notebook: default
---



```{r}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE,
                      fig.align = "center", 
                      fig.width = 6, 
                      fig.height = 5,
                      out.height = "40%")
set.seed(12314159)
library(loon.data)
library(loon)
library(gridExtra)

codeDirectory <- "../../img"
imageDirectory <- "./img"
dataDirectory <- "./data"
path_concat <- function(path1, ..., sep="/") paste(path1, ..., sep = sep)

```

# (3)

## The full data set is then read in as 
```{r}
labData <- read.csv("labData.csv")
```

# (A)

```{r}
observational <- labData[labData$type == "observational", ]
```


# (4)

# (B)

```{r}

plot(observational$x, observational$y, xlim = c(0, 30), ylim = c(0, 30), pch = 19, col = adjustcolor("black", 0.3), main = "(x,y) pairs from all of the observational Data", xlab = "x-values", ylab = "y-values")

Bmod <- lm(y~x, observational)
abline(Bmod)
#print(Bmod$coefficients)
print("Slope Estimate is ")
print(Bmod$coefficients[2])
```

# (C)

## (i)

```{r}
plot(observational$x, Bmod$residuals, xlim = c(0, 30), abline(h = 0), main = "Residuals against the X values", xlab = "x-values", ylab = "residuals")
```

## (ii)

### From the above residual plot, we can infer that the range of the residuals increases when the value of x increases. I would say that the measuring system is baised towards the low X values.


# (D)

## (i)

```{r warning=FALSE}
fit <- array()
estimates <- c()

for (value in 1:18){
  fit[value] <- lm(y~x, observational[observational$team == value, ])
  estimates <- c(estimates, fit[[value]][2])
}
print("Average of estimated slopes is ")
print(mean(estimates))
```


## (ii)


## (iii)

```{r}

hist(estimates, xlim = c(-1, 1), col = "lightgrey", main = "Individual slope cofficient estimates", xlab = "slope estimates")
abline(v = mean(estimates), col = "red", lty = 2)
abline(v = 0.5, col = "blue", lty = 2)

```


# (5)

# (E)
```{r}
print(estimates)
```

###  From the observational study and seeing the estimate values, it is evident that the estimates lies between -0.6171 and -0.2885. On the contrary the true value is 0.5. Hence i would conclude that the quality of slope estimates is bad.  

# (F)

### Since Z is a lurking variable for the above problem it is clear that, it has a fixed value which fixes (hyperplane). This has a effect on the values of y. Due to this the y is being restricted to take the values. 