---
title: "EDA-Assignment1 Ram Yoogesh"
output:
  html_document:
    df_print: paged
  html_notebook: default
---

## Question 2

## Loading the CSV data 

```{r}
path_concat <- function(path1, path2, sep="/") {
paste(path1, path2, sep = sep)
}
dataDirectory = "/home/yoogesh/Desktop/stat847assignments/data/"
data <- read.csv(file = path_concat(dataDirectory, "class_data.csv"))

```

## Previwing the content

```{r}
#View(data)
data
str(data)
names(data)
```


## 1 
## a)

### Hypothesis is false if the card a shows the person age less than 19. Hypothesis is true for card b - reason is that it shows the picture of an beverage , so it doesn't matter what age the person is. Hypothesis is false for card c if it shows alcohol else its true. Card d is like b, that is it shows the age is greater than 9, so we dont care about the drink

## b)
### Cards A and C 

## c)
   \newline
   $\binom{2}{2}$ $/$ $\binom{4}{2}$
   $=$ $\frac{1}{6}$

## d)
```{r}
tail(names(sort(table(data$green_card1))), 1)

gc <-tail(names(sort(table(data$green_card2))), 1)
gc

length(data[data[,4] == gc ,1])
```
## Its correct

## e)
## (i)
```{r}
correct_g1 <- ifelse(data$green_card1 =="a" | data$green_card1 =="c", "TRUE", "FALSE")
correct_g2 <- ifelse(data$green_card2 =="c" | data$green_card2 =="a", "TRUE", "FALSE")
```

## (ii)
```{r}
propo <- prop.table(table(correct_g1))
propo ["TRUE"]
choose(2,1)/choose(4,1)
```

## (iii)
```{r}
tot  <- sum(correct_g1== "TRUE" & correct_g2 == "TRUE")
NROW(correct_g1)
tot/NROW(correct_g1)
choose(2,2)/choose(4,2)
```

## (iv) 
```{r}
r = which(correct_g1 =="TRUE")
sum(correct_g2[r]=="TRUE")/length(r)
choose(2,1)/choose(4,2)
```

## (V)
```{r}
prop_correct_green <- ifelse(correct_g1 =="TRUE" & correct_g2 =="TRUE", 1, ifelse((correct_g1 =="TRUE" & correct_g2 =="FALSE") | (correct_g1 =="FALSE" & correct_g2 =="TRUE") , 0.5, 0))
```

```{r}
hist(prop_correct_green)
abline( v = mean(prop_correct_green), lwd=3, lty=2, col="green")
abline( v = median(prop_correct_green), lwd=3, lty=2, col="green")
```

## (vi)
### Overall people did well. 


## 2)

## a)
### Hypothesis is true if card a reveals an even number else its false
### Hypothesis is true since card b has a vowel
### Hypothesis is true for card c since it shows an even number
### Hypothesis is true if card d doesn't reveal a vowel else its false

## b)
### Cards A and D are correct 

## c)
```{r}
tail(names(sort(table(data$red_card1))), 1)
r <- tail(names(sort(table(data$red_card2))), 1)
length(data[data[,6] ==r ,1])
r
```
### Its wrong

## d)
##(i)

```{r}
correct_r1 <- ifelse (data$red_card1 =="a" | data$red_card1 =="d", "TRUE", "FALSE")
correct_r2 <- ifelse (data$red_card2 =="d" | data$red_card2 =="a", "TRUE", "FALSE")
```


## (ii)
```{r}
propo <- prop.table (table(correct_r1))
propo ["TRUE"]
choose(2,1)/choose(4,1)
```

## (iii)
```{r}
tot <- sum (correct_r1== "TRUE" & correct_r2 == "TRUE")
NROW (correct_r1)
tot/NROW (correct_r1)
choose(2,2)/choose(4,2)
```

## (iv) 
```{r}
r = which(correct_r1 =="TRUE")
sum (correct_r2[r]=="TRUE")/length(r)
choose(2,1)/choose(4,2)
```

## (v)
```{r}
prop_correct_red <- ifelse(correct_r1 =="TRUE" & correct_r2 =="TRUE", 1, ifelse((correct_r1 =="TRUE" & correct_r2 =="FALSE") | (correct_r1 =="FALSE" & correct_r2 =="TRUE") , 0.5, 0))
hist(prop_correct_red)
abline( v = mean(prop_correct_red), lwd=3, lty=2, col="green")
abline( v = median(prop_correct_red), lwd=3, lty=2, col="green")
```
## (vi)
### Overall people performed badly



## 3)
## a)
### Both are same since the probablity is same. 
 
## c)
## (i)
 pr(p=0)=
 $\binom{2}{1}$ $/$ $\binom{4}{2}$
   $=$ $\frac{1}{3}$
   
   pr(p=0.5)=
 $\binom{2}{1}$ $x$ $\binom{2}{1}$ $/$ $\binom{4}{2}$
   $=$ $\frac{2}{3}$
   
   pr(p=1)=
 $\binom{2}{1}$  $/$ $\binom{4}{2}$
   $=$ $\frac{1}{3}$
   
## (ii)

$E(p)$ $=$ $0*1/3$ $+$ $1*1/3$ $+$ $0.5*2/3$
$=$ $0.667$


   
   
